---
title: "STAT 545A Homework 2"
author: "Tian Gao"
date: "2018/9/20"
output: github_document
---


# Bring rectangular data in
```{r}
library(tidyverse)
library(gapminder)
```

# Smell test the data
## Is it a data frame, a matrix, a vector, a list?
```{r}
class(gapminder)

```
* This is a data frame

## what is its class?
* Its classes are tbl_df, tbl, data.frame

## How many variables/columns?
```{r}
ncol(gapminder)
```

## How many rows/observations?
```{r}
nrow(gapminder)
```
## Can you get these facts about "extent" or "size" in more than one way? Can you imagine different functions being useful in different contexts?
```{r}
dim(gapminder)
```

## What data type is each variable?
```{r}
str(gapminder)

```
* We can see that there are six variables:
* "country" is Factor
* "continent" is Factor
* "year" is int
* "lifeExp" is num
* "pop" is int
* "gdpPercap" is num
 
 
# Explore individual variables
## What are possible values of each variables? 
```{r}
unique(gapminder$year)
range(gapminder$year)
```
* Above are the unique values and range of "year" variable
```{r}
unique(gapminder$continent)
```
* Above are the unique values of "continent" variable.

## What values are typical? What's the spread? What's the distribution? 
```{r}
frequency_year = table(gapminder$year)
frequency_continent = table(gapminder$continent)
frequency_year
frequency_continent
```
* We can see that, from the aspect of year, the sample is equally distributed, every year appeard 142 times
* From the aspect of continent, the typical value is "Africa"
```{r}
summary(gapminder$year)
summary(gapminder$continent)
```

# Explore various plot types
## Make a few plots, probably of the same variable you chose to characterize numerically. You can use the plot types we went over in class (cm006) to get an idea of what you¡¯d like to make. Try to explore more than one plot type.
* First I'll explore the year and LifeExp
```{r}
# year as x axis and lifeExp as y axis
ggplot(gapminder,aes(x=year, y=lifeExp))+
  # scale y axis by log10
  scale_y_log10()+
  # make a scatterplot
  geom_point()
```
* Here is the jitter version
```{r}
ggplot(gapminder, aes(x=year, y=pop))+
  scale_y_log10()+
  # make a jitter plot
  geom_jitter()
```
* Next I'll focus on canada and plot the changes in canadian's lifeExp against year.
```{r}
# filter the country to see only Canada
  filter(gapminder, country == "Canada") %>%
  ggplot(aes(x=year, y=lifeExp))+
  scale_y_log10()+
# plot both point and line to see the trend
  geom_point()+
  geom_line()

```
* Next I'll plot the histgram, densityplot and frequency polygon of population 
* A histogram of "pop"
```{r}
ggplot(gapminder, aes(x=pop)) + 
  geom_histogram(bins=100)+
  scale_x_log10()
```
* A densityplot of "pop"
```{r}
ggplot(gapminder, aes(x = pop))+
  geom_density(fill="pink")+
  scale_x_log10()
```
* I'll then plot both of the histogram and densityplot in one graph.
```{r}
ggplot(gapminder, aes(x=pop)) + 
  scale_x_log10()+
  geom_histogram(bins=100, aes(y=..density..))+
  geom_density()
```
* A frequency polygon of "pop"

```{r}
ggplot(gapminder, aes(x = pop))+
  scale_x_log10()+
  geom_freqpoly(bins=100)
```
* Next I'll use boxplots to show the population of different continents
```{r}
# select population and continent
select(gapminder, pop, continent) %>%
  ggplot(aes(x = continent, y=pop)) + 
  scale_y_log10()+
# make a boxplot
  geom_boxplot()
```


# But I want to do more
## Evaluate this code and describe the result
```{r}
filter(gapminder, country == c("Rwanda", "Afghanistan"))
```

* Evaluate this code and describe the result. Presumably the analyst¡¯s intent was to get the data for Rwanda and Afghanistan. Did they succeed? Why or why not? If not, what is the correct way to do this?
* I think we should use %in% instead of ==, because the original intent is to get data either of Afghanistan or Rwanda, but not both. The following is the right version.

```{r}
filter(gapminder, country %in% c("Rwanda", "Afghanistan"))
```


*Explore summarise() and group_by()
```{r}
group_by(gapminder, continent) %>%
summarize(
    avg_lifeExp = mean(lifeExp),
    average_gpd = mean(gdpPercap),
  
)
```

